An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-27 draft.tex
Starting Experimental Setup

\usepackage{lineno,hyperref}
\usepackage{rotating}
\usepackage{algpseudocode,algorithm}
\usepackage{hyperref}

@@

\journal{Future Generation Computer Systems}
\begin{document}

\begin{frontmatter}

@@

The COCO framework offers several tools to compare the performance of
algorithms, generating data sets, tables, and reports for an experiment. There
is a repository of more than 200 results for the noiseless BOBB testbed,
collected from BBOB workshops and special sessions between the years 2009 and
2019 \url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}. The
EvoSwarm application includes an adapted version of the noiseless BOBB testbed,
compatible with the scripts of the framework, to compare with other algorithms
in the repository.

To validate the heterogeneous multi-population setting, we compare between
homogeneous and an ensemble of multi-populations, using Genetic Algorithms
(GAs) and Particle Swarm Optimization (PSO). For the GA implementation, we used
the DEAP library \cite{fortin2012deap} and for the PSO, the EvoloPy library
\cite{faris2016evolopy}. Both Python libraries are open-source. In EvoSwarm 
these two algorithms are implemented as stateless functions.

We show the parameters for each algorithm, GA in Table \ref{tab:GAparams} and
for PSO in Table \ref{tab:PSOparams}. We obtained these parameters in a
previous work \cite{}. To obtain the parameters, we tested first on the
Rastrigin separable function with five dimensions. After about fifteen
experiments, the most challenging targets were achieved for this particular
function. We tested again with functions one to three, and after obtaining
favorable results, the PSO and GA algorithm parameters were set.In the GA, we
randomly set the mutation and crossover probabilities to have more
heterogeneous workers; in these parameters, the range of values is specified.

\begin{table}
  \small
  \caption{ DEAP GA EvoWorker Parameters }
  \label{tab:GAparams} 
  \centering
  \small
  \begin{tabular}{|l|c|}
    \hline
    Selection & Tournament size=12\\ \hline
    Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05  \\ \hline
    Mutation Probability & [0.1,0.3]  \\ \hline
    Crossover & Two Point  \\ \hline
    Crossover Probability& [0.2,0.6]  \\ \hline
  \end{tabular}
\end{table}

@@

\begin{table}
    \small
    \caption{10 Populations experiment parameters
    }
    \label{tab:params:workers}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{5 Populations experiment parameters
    }
    \label{tab:params:workers}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

>>>>

\usepackage{lineno,hyperref}
\usepackage{rotating}
\usepackage{algpseudocode,algorithm}
\usepackage{hyperref}
\usepackage{listings}

@@

\journal{Future Generation Computer Systems}
\begin{document}
\lstset{language=Python} 
\begin{frontmatter}

@@

The COCO framework offers several tools to compare the performance of
algorithms, generating data sets, tables, and reports for an experiment. There
is a repository\footnote{\url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}} 
of more than 200 results for the noiseless BOBB testbed, collected from 
BBOB workshops and special sessions between the years 2009 and2019. The
EvoSwarm application includes an adapted version of the noiseless BOBB testbed,
compatible with the scripts of the framework, to compare with other algorithms
in the repository.

To test the heterogeneous multi-population capabilities,  we compare the
performance between a homogeneous and an ensemble of multi-populations, using
Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). For the GA
implementation, we used the DEAP library \cite{fortin2012deap} and for the  PSO,
the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
open-source. In EvoSwarm these two algorithms are implemented as stateless
functions.

Next, we show the parameters for each algorithm, in Table \ref{tab:GAparams} 
for the GA and in Table \ref{tab:PSOparams} for PSO. We obtained these parameters 
following the same method as in a previous work \cite{garcia2017benchmarking}. 
To obtain the parameters, we tested first on the
Rastrigin separable function with five dimensions. After about fifteen
experiments, the most challenging targets were achieved for this particular
function. We tested again with functions one to three, and after obtaining
favorable results, the PSO and GA algorithm parameters were set.In the GA, we
randomly set the mutation and crossover probabilities to have more
heterogeneous workers; in these parameters, the range of values is specified.

\begin{table}
  \small
  \caption{ DEAP GA EvoWorker Parameters }
  \label{tab:GAparams} 
  \centering
  \small
  \begin{tabular}{|l|c|}
    \hline
    Selection & Tournament size=12                            \\ \hline
    Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05   \\ \hline
    Mutation Probability & [0.1,0.3]                          \\ \hline
    Crossover & Two Point                                     \\ \hline
    Crossover Probability  & [0.2,0.6]                          \\ \hline
  \end{tabular}
\end{table}

@@

% Experiment configuration

\begin{table}
    \small
    \caption{ Experiment configuration example 
    }
    \label{tab:params}
    \centering
    \small
    \begin{tabular}{|l|l|l|l|l|}
      \hline
      \multicolumn{3}{|l|}{Parameter}                   & Type             & Example         \\ \hline
      \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
      \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
      \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
      \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
      \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
      \multicolumn{3}{|l|}{Mutation  Probability Rrange} & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
      \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
      Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
              10              & 50      & 140                 &      5                 & 30                \\ \hline
              20              & 66      & 200                 &      5                 & 30               \\ \hline
              40              & 80      & 250                 &      5                 & 40                \\ \hline
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{10 Populations experiment parameters
    }
    \label{tab:params:10}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{5 Populations experiment parameters
    }
    \label{tab:params:5}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}
